---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nphawkes

<!-- badges: start -->
<!-- badges: end -->

The nphawkes package can be used to study Hawkes processes using nonparametric procedures. Data may be applied to the Model Independent Stochastic Declustering (MISD) algorithm to probabilistically classify events as background or triggered events. The conditional intensity of the process can be calculated, and residual analysis may be exectued via thinning, superpositioning, or super-thinning. This package may also be used to create plots such as histogram estimators of the triggering functions, line plots of the conditional intensity, and histograms of the residual process. Temporal and spatio-temporal Hawkes processes can be used, along with marked versions of each. 

## Installation

<!-- You can install the released version of nphawkes from [CRAN](https://CRAN.R-project.org) with: -->

You can install the resleased version of nphawkes from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("boydpe/nphawkes")

```
## Example

We will demonstrate the use of the package with simulated earthquake data available within the package. 

```{r example}
library(nphawkes)
load(file = "data/eq.RData")
head(quake)
```

We will first employ the MISD algorithm to decluster the earthquake catalog using time, space, and marks (magnitude). The temporal bins (time_breaks) will be defined, in terms of days, and we will use roughly uniform bins for space and marks by setting the `space_quantile` and `mark_quantile` inputs equal to `TRUE`. 

```{r}
out = nph(dates = quake$date, ref_date = "2010-01-01",
          lat = quake$lat, lon = quake$lon,
          marks = quake$m,
          time_breaks = c(0, 1, 3, 5, 7, 14, 31, 1500),
          #mark_breaks = c(3, 3.1, 3.25, 3.4, 6),
          space_breaks = c(0, 5, 10, 20, 50, 100),
          time_quantile = FALSE,
          space_quantile = FALSE,
          mark_quantile = TRUE,
          time_of_day = quake$time)

out1 = nph(dates = quake$date, ref_date = "2010-01-01",
          #lat = quake$lat, lon = quake$lon,
          marks = quake$m, 
          time_breaks = c(0, 1, 3, 5, 7, 14, 31, 1500),
          #mark_breaks = c(3, 3.1, 3.25, 3.4, 6),
          time_quantile = FALSE,
          #space_quantile = TRUE,
          mark_quantile = TRUE,
          time_of_day = quake$time)

out2 = nph(dates = quake$date, ref_date = "2010-01-01",
          lat = quake$lat, lon = quake$lon,
          #marks = quake$m,
          time_breaks = c(0, 1, 3, 5, 7, 14, 31, 1500),
          #mark_breaks = c(3, 3.1, 3.25, 3.4, 6),
          time_quantile = FALSE,
          space_quantile = TRUE,
          #mark_quantile = TRUE,
          time_of_day = quake$time)

out4 = nph(dates = quake$date, ref_date = "2010-01-01",
          lat = quake$lat, lon = quake$lon,
          marks = quake$m,
          time_breaks = c(0, 5, 7, 14, 31, 1500),
          #mark_breaks = c(3, 3.1, 3.25, 3.4, 6),
          #space_breaks = c(0, 5, 10, 20, 50, 100),
          time_quantile = FALSE,
          space_quantile = TRUE,
          mark_quantile = TRUE,
          h_length = 4,
          k_length = 4,
          #mark_quantile = FALSE,
          time_of_day = quake$time)

```

We can view the background rate and percentage of mass that lies on the diagonal of the resulting probability matrix. 

```{r}
out$br
out$perc_diag
```

We can then inspect histogram estimators of the triggering functions. 

```{r}
trig_plots(model = out4, g_max = 32,
           k_max = 4, k_min = 3,  
           h_max = 22)
```

We'll now calculate the conditional intensity of the process. 

```{r}
ci = cond_int(model = out)
head(ci)
```

We can perfrom super-thinning as a residual method, thinning events in areas of high conditional intenstiy and superpositioning events in areas of low intensity. We'll use the median value of the conditional intensity as the threshhold value. We'll use the `sim_grid = TRUE` argument since the geographical coordinates are simulated within the $[0\times1]\times[0\times1]$ grid. 
 

```{r}
st = super_thin(K = "median_ci", model = out,
                method = "superthin", sim_grid = TRUE)
```

