---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nphawkes

<!-- badges: start -->
<!-- badges: end -->

The nphawkes package can be used to study Hawkes processes using nonparametric procedures. Data may be applied to the Model Independent Stochastic Declustering (MISD) algorithm to probabilistically classify events as background or triggered events. The conditional intensity of the process can be calculated, and residual analysis may be exectued via thinning, superpositioning, or super-thinning. This package may also be used to create plots such as histogram estimators of the triggering functions, line plots of the conditional intensity, and histograms of the residual process. Temporal and spatio-temporal Hawkes processes can be used, along with marked versions of each. 

## Installation

<!-- You can install the released version of nphawkes from [CRAN](https://CRAN.R-project.org) with: -->

You can install the resleased version of nphawkes from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("boydpe/nphawkes")

```
## Example

We will demonstrate the use of the package with simulated earthquake data available within the package.  

```{r example}
library(nphawkes)
load(file = "data/quake.RData")
head(quake)
```

We will first employ the MISD algorithm to decluster the earthquake catalog using time and marks (magnitude). The temporal bins (time_breaks) will be defined, in terms of days, and we will use roughly uniform bins for marks by setting the `mark_quantile` inputs equal to `TRUE`. We'll set out reference date to be January 1, 2010, the date that our simulated catalog began. 

```{r}
t1 = c(0, 0.001, 0.01, 0.1, 1, 5, 10, 100, 1500)
t2 = c(0,2,5,10, 31, 1500)
s1 = c(0, 0.01, 0.1, 1, 10, 100)
s2 = c(0, 0.001, 0.01, 0.1, 1, 5, 10, 50, 100)
s3 = c(0, 0.1, 1, 5, 10, 25, 50, 100)
s4 = c(0, 0.001, 0.01, 0.25, 0.50, 0.75, 1, 2.5, 5, 10, 25, 50, 100)
out = nph(dates = quake$date, ref_date = "2010-01-01",
          marks = quake$m, 
          lat = quake$lat,
          lon = quake$lon,
          time_breaks = t1,
          space_breaks = s2,
          space_quantile = FALSE,
          time_quantile = FALSE,
          mark_quantile = TRUE,
          time_of_day = quake$time)

# good luck so far with t1 and mark quantile

```

We can view the background rate and percentage of mass that lies on the diagonal of the resulting probability matrix. 

```{r}
out$br
out$perc_diag
```

We can then inspect histogram estimators of the triggering functions. The triggering plots include standard error estimates, which can be separately calculated using `se_bars(model)`. 

```{r}
trig_plots(model = out, 
           g_xlim = c(0,32),
           k_xlim = c(3,4))
```

We'll now calculate the conditional intensity of the process. 

```{r}
ci = cond_int(model = out)
head(ci)
```

We can perfrom super-thinning as a residual method, thinning events in areas of high conditional intenstiy and superpositioning events in areas of low intensity. We'll use the median value of the conditional intensity as the threshhold value. 
<!-- We'll use the `sim_grid = TRUE` argument since the geographical coordinates are simulated within the $`[0\times1]\times[0\times1]`$ grid.  -->
 

```{r}
st = super_thin(K = "median_ci", model = out,
                method = "superthin")
head(st)
```

After super-thinning, we can view a histogram of the residual process to assess the resulting distribution.

```{r}
ci_hist(st)
```

We can view the conditional intensity of the process, plotted against the frequency of events. The conditional intensity is 

```{r}
ci_plot(model = out, superthin = st)
```

Finally, we can inspect a plot that shows where events are superposed and where events are thinned. This tiered plot shows, over time, events that were not thinned (retained), superposed events, and thinned events. 

```{r}
st_plot(st)
```


```{r, include=FALSE, echo = FALSE, eval = FALSE}
out = nph(dates = quake$date, ref_date = "2010-01-01",
          lat = quake$lat, lon = quake$lon,
          marks = quake$m,
          time_breaks = c(0, 3, 10, 31, 1500),
          mark_breaks = c(3, 3.1, 3.25, 3.4, 6),
          space_breaks = c(0, 10, 25, 50, 100),
          # time_quantile = TRUE,
          # space_quantile = TRUE,
          # mark_quantile = TRUE,
          time_of_day = quake$time)
```

