---
title: "Nonparametric Hawkes Processes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nphawkes)
```

```{r, include=FALSE}
# Data Setup
# quake = read.csv("~/OSU/MS/eq_test1.csv")
# refd = "1/1/2010"
# full_time = lubridate::mdy(refd) + lubridate::duration(quake$t, "days")
# quake$date = lubridate::date(full_time)
# quake$time = paste(lubridate::hour(full_time),lubridate::minute(full_time),round(lubridate::second(full_time)), sep=":")

```

In this document, we will demonstrate the use of the `nphawkes` package. We will use the `eq.RData` file, found in this package. 

This file contains 471 simulated earthquakes and contains the location in latitude and longitude (in [0,1] $x$ [0,1] unit square), date, time of day, and magnitude ($m$) of each event. 

```{r, echo=FALSE}
quake = data("eq.RData")
quake = data("eq")
quake = data(eq)
head(quake)
```

We will first employ the Model Independent Stochastic Declustering algorithm via the `nph()` function. This function must be executed at least in terms of a temporal Hawkes process, but it may also be analyzed as a spataio-temporal or marked spatio-temporal Hawkes process. Provided the given information (date, spatial location, and marks), we will use all three triggering components. 

```{r}
out = nph(dates = quake$date, lat = quake$lat, lon = quake$lon,
          marks = quake$m, g_bins = c(0, 7, 31, 180, 365, 1000), 
          h_quantile = TRUE, k_quantile = TRUE, 
          time_unit = "day", time_of_day = quake$time)
```




